Sketch Recognition and Understanding

### 2009-04-28 draft/chapterSurvey.tex
before edit in swarm
before fager

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize independent symbols. Each symbol is segmented and hierarchical features are computed.  Features that are representing the global description of the shape are also generated and appended to the hierarchical features. Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

\chapter{Particle Swarm Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swarm Optimization Review}
\label{sec:ParticleSwramOptimizationReview}
%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
% start with swarm in general  .. say about flock , ant and usin in image processing 
% then speeak about PSO and  its firstly generated   (see article)
%mention that is best than ga (see thesis)

@@

Image processing is one of the disciplines that used Swarm Intelligence. Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}. Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}. PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.

\section{Discrete Particle Swarm Algorithm}
\label{sec:ParticleSwarmAlgorithm}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}. Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
\label{eq:Swarm}
\end{equation}
 where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}. 
 \begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i} 
\label{eq:descrite}
\end{array}\}
\end{equation}
 where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

>>>>

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize independent symbols. Each symbol is segmented and hierarchical features are computed.  Features that are representing the global description of the shape are also generated and appended to the hierarchical features. Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms. 
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
\chapter{Particle Swarm Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swarm Optimization Review}
\label{sec:ParticleSwramOptimizationReview}

@@

The underliny research in swarm intelligence led to optimization algorithm that simulate the animal social behvaiors. 
Ant Colony Optimization (ACO) algoirithm simulate how ants search for optimize path for food \cite{ACOFirst}. Real ants leave evidence of their exploriation while they are searching for food. This trail is used by other ants to locate shortest distance between the colony and food location. ACO use similar type of cooperation between agents to find optimal solution.  The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.  Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).

   The next section discribes in details the Particle Swarm Optimization and its advantages and disatvantages. 
% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO). The algorithm simulate the moving  Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm. 
 \subsection*{Particle Swarm Algorithm}
 In optimization problems, the algorithm attempts to find a feasable solution which satisfy all constrain of this problem. An optimal solution is the best feasable solution in the solution space. Several methods were used to solve optimiation problems. Stochastic search is based on the assumption that good solution are near each other. Algoirthms like simluating aneeling and hill climbing used stochastic to find best solution. The problem in these methods is they may be traped into a local best and fail to find the global best solution.

  Evolutionary algorithms (e.g. Genatic Algoirthms) explores solution using a population which is evalutated using fitness functions. The algorithms initially start with a randomly choosen population which is used to generate new population using various operators (e.g. cross over, mutation... ).  After each iteration the population is evaluated using a fitness function and only the strongest are used to generate new population. The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not alway converge to a global best solution\cite{PSOpattern}.

 Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space. A fitness function is used to evaluate the goodness of each particle in solution space. But unlike evolutionly algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fitest. Each particle saves a history of the best location in the swarm and its own best location. After each iteration, the particles moves to a new location based on the global best location and it own history of best location. This use of pervious particle locations gives PSO advantage over Evolutionary algoirthms where no similar information is used.

 PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}. In general the PSO is a continous optimization problem but expriments showed it performed as GA is discrite and binary optimizaition problems. The comparison between PSO and Genetic algorithm was perofrmed in \cite{PSOGAComp} and  proved that PSO generate faster, better solutions than GA specially in problems with high dimensinalty.

 A hybrid of PSO and other algorithms was investigated in various systems. \cite{} used a hybird of GA and PSO to breed solutions using GA then feed them to PSO. They domenstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.

 \subsection*{PSO neighbourhood topology}  
In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle. Other forms of cooperation and organizatoin were suggested in various systems.  \cite{} suggested to use a grid like topology where particle are connected to four other particles sround it. A ring  and star topology were widley used in \cite{}. In \cite{hiericalPSO}, Janson and Middendorf proposed to use a dynamic and hierrchinal neighbourhood structure. They showed that using such structure helps in mentainaing diversity over the solution space in a dynamic problem. % a PSO using a dynamic and hierarchical neighbourhood structure to handle dynamic optimization problems. They demonstrated that such a structure is useful for maintaining some particle diversity in a dynamic environment.

 Other forms of cooperations were also investigated in \cite{facotrsSwarm} where the problem is divided among particles and each particle solves part of the problem. \cite{facotrsSwarm} discuss different behvaiorus that governs cooperation particles in the swarm system.  %were used  There is various toplogyies used in connecting a swarm in PSO. The most used is the totaly connected, star and ring topology.

\subsection*{Drawbacks}
The main drawback of the particle swarm algorithm is the premanture coverge problem. This happens when global best cannot be improved after a few swarm iterations. This is because the use of global best makes pso converge to one place which is between local best and global best.  Another resons is because of the high amount of information between particle makes the solution less diversy which limit the explored solution space. This can be solved by using high interita $w$ or using another improvement that generate new random populations.

Another problem of PSO is that performance is dependent on the problem and the PSO paramter selected for this specific problem. This means that any change in the PSO paramter change the efficency of PSO solution.

\subsection*{Improvements to PSO}

change so that the optima can still be tracked.
One of the early works on using PSO for dynamic optimization
was by Eberhart and Shi in [13], where they used
an inertia weight version of PSO to track the optimum of
a 3-dimensional unimodal parabolic function which changes
its maxima every 100 iterations. It was found under certain
circumstances that the PSO’s performance was comparable
or better than that of previously published evolutionary algorithms
[1, 2].
For detection, Carlisle and Dozier used a sentry particle
which is randomly chosen at each iteration [10]. The sentry
particle gets evaluated before each iteration and compares
its fitness with its previous fitness value. If the two values
are different, indicating the environment has changed, ie.,
the optimum has moved, then the whole population gets
alerted and several possible responses can then be triggered.
Hu and Eberhart proposed to re-evaluate  pg and a second
best particle to detect if a change has occurred.
Various response strategies have also been proposed. Carlisle
and Dozier proposed to re-evaluate all pi of all particles when
a change has been detected [11]. If the re-evaluated pi is less
fit than its current position xi, then this pi is replaced by
its associated xi, because only these two evaluations can
be considered up-to-date. Hu and Eberhart studied the effects
of re-randomizing various proportions of the swarm to
maintain some degree of diversity in order to better track
the optima after a change [14]. However, this approach suffers
from possible information loss since the re-randomized
portion of the population does not retain any information
that might be useful from the past iterations.
In order to maintain better particle diversity throughout
a run, Blackwell and Bentley introduced charged swarms
where mutually repelling charged particles orbit a nucleus of
neutral particles (conventional PSO particles) [3]. Whereas
the charged particles allow the swarm to better adapt to
changes in the environment, the neutral particles play the
role of continuing to converge towards the optimum in finer
details.
Inspired by multi-population EA approaches such as the
multinational GA [22] and self-organizing scouts [7], Blackwell
and Branke proposed an interacting multi-swarm PSO
as a further improvement to the charged swarms [4]. The
multi-swarm PSO aims at maintaining multiple swarm populations
on different peaks. Multiple swarms are prevented
from converging to the same optimum by randomizing the
worse of two swarms that come too close. The multi-swarm
PSO also replaces the charged particles by quantum particles
whose position is solely based on a probability function
centered around the swarm attractor. The resulting multiquantum
swarms outperform charged and standard PSO on
the moving peaks function. This multi-swarm approach is
particularly attractive because of its improved adaptability
in a more complex multimodal dynamic environment where
multiple peaks exist and need to be tracked.
With a similar aim to locate and track multiple peaks
in a dynamic environment, Parrott and Li in [19, 20] proposed
a species-based PSO (SPSO) incorporating a speciation
algorithm first proposed by Petrowski [21]. The SPSO
uses a local “species seed” which provides the local  pg to
particles whose positions in the search space are within a
user-specified radius of the seed. This encourages swarms
to converge onto multiple local optima instead of a single
global optimum, hence developing multiple sub-populations
in parallel. In addition, the dynamic SPSO uses a parameter
pmax to limit the number of particles allowed in a species
(or swarm), with the excess particles reinitialized at random
positions in the total search space.

\subsection*{Particle Swarm Applications}
\label{sec:SwarmApp}
 Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in various systems. The algorithm was used to train Neural network in \cite{} and proved to perform better than GA algoirthm. In Image processing decipline PSO was used an optimization algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}. PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.

\section{Discrete Particle Swarm Algorithm}
\label{sec:ParticleSwarmAlgorithm}

  %It simulate the  % which has the best or  select a set of paramters that satisfy all constrains

\subsection*{Discrite Particle Swarm Equations(DPSO)}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
where $p_{ij}$ represent the $jth$ dimention in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimention in the $ith$ particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} W  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
\label{eq:Swarm}
\end{equation}
 where $W$ is the inertia weight paramter which controls the trade off between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration paramters.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space. The velocity is bounded to maximum of $V_{max}$ using Equation\ref{eq:vmax} to prevent exploding which leads to premature convergence. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}. 
 \begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i} 
\label{eq:descrite}
\end{array}\}
\end{equation} where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable. 
  \begin{equation}
V_{ij}  = \frac{{V_{ij} }}{{2V_{\max } }} + 0.5
\label{eq:vmax}
\end{equation} where $V_{\max }$ is the velocity threshold.

\subsection*{The  Algorithm}
%\label{sec:ParticleSwarmAlgorithm} 
\begin{verbatim}
 Initialize 
    Generate N particles with M dimentions (P_i...P_N) 
    Generate velocities for each particle for M dimention randomly. 
 Repeat till Maximum number of iteration OR Fitness 
         For each particle P_i do the following 
                Compute fitness  
                Determine Lbest  
                Check if need to change Gbest. 
                Compute the Velocity for each dimention using eq. 
                Check if maximum Velocity reach and limit it 
                update particle values 
          end . 
        Improve all particle positions 
  end

  Final solution reach in Gbest.

\end{verbatim}
